<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{alm vignette}
-->

```{r, eval=TRUE, echo=FALSE}
knitr::opts_chunk$set(
  fig.width=6, 
	fig.height=5, 
	fig.pos="h",
	message=FALSE, 
	warning=FALSE, 
	comment="#>", 
	cache=TRUE
)
```

alm tutorial
============

The `alm` package is a set of functions to access article level metrics via a RESTful API from the Rails app `Lagotto` created by the Public Library of Science (PLOS). `Lagotto` is being used by PLOS, and a number of other publishers:

* PLOS (all their journals) at [http://alm.plos.org/](http://alm.plos.org/)
* PLOS test server at [http://labs.lagotto.io/](http://labs.lagotto.io/)
* Copernicus (seems to be down for now)
* Public Knowledge Project (PKP) at [http://pkp-alm.lib.sfu.ca/](http://pkp-alm.lib.sfu.ca/)
* Crossref at [http://det.labs.crossref.org/](http://det.labs.crossref.org/)
* eLife at [http://lagotto.svr.elifesciences.org/](http://lagotto.svr.elifesciences.org/)
* Pensoft at [http://alm.pensoft.net:81/](http://alm.pensoft.net:81/)
* Making Dat Count at [http://dlm.plos.org](http://dlm.plos.org)

A good place to look for the status of various installations of Lagotto is this status page: http://articlemetrics.github.io/status/ (which also includes what version of Lagotto each is running)

## Help with Lagotto

Lagotto has a nice support site at [http://discuss.lagotto.io/](http://discuss.lagotto.io/) for any questions about it.

## What is an article level metric?

Glad you asked. The canonical URL for this is perhaps [altmetrics.org](http://altmetrics.org/manifesto/). Basically it is a metric that measures something about an article. This is in stark contrast to journal level metrics, like the [Journal Impact Factor](http://www.wikiwand.com/en/Impact_factor).

## Are there other altmetrics data providers?

Yes indeedy, but see notes

+ [ImpactStory](http://impactstory.org/) - No open data (except for your own data)
+ [Altmetric.com](http://altmetric.com/) - Some open data
+ [PlumAnalytics](http://www.plumanalytics.com/) - No open data

## Authentication

You only need an API key for the publishers PKP and Pensoft. You can set the key in your options just for the current session by executing `options(PlosApiKey = "YOUalmAPIKEY")`, or pass in to each function call with the `key` parameter, or save in your `.Rprofile` file.

## URLs

The default URL is set for the PLOS data sources: http://alm.plos.org/api/v5/articles
You can change this URL. For example, if you want to get data from the Crossref instance, set the `api_url` parameter to http://alm.labs.crossref.org/api/v5/articles

## Other languages

If R is not your thing, there are Lagotto clients in development for [Ruby](https://github.com/articlemetrics/lagotto-rb) and [Python](https://github.com/articlemetrics/pyalm).

## Help 

If you are having trouble with this R package, contact [the maintainer, Scott](mailto:myrmecocystus@gmail.com). If you are having trouble with the API itself, there is a newish discussion forum for the Lagotto service at [http://discuss.lagotto.io/](http://discuss.lagotto.io/). 

## Install and load

Stable version from CRAN

```{r eval=FALSE}
install.packages("alm")
```

Or development version from GitHub

```{r install, eval=FALSE}
install.packages("devtools")
devtools::install_github("ropensci/alm")
```

Load `alm`

```{r}
library("alm")
```

## Get data

```{r alm1}
alm_ids(doi="10.1371/journal.pone.0029797")
```

## Details for a single DOI

```{r alm2}
out <- alm_ids(doi='10.1371/journal.pone.0029797', info='detail')
```

info

```{r}
out$data$info
```

sign posts

```{r}
out$data$signposts
```

totals

```{r}
out$data$totals
```

history

```{r}
out$data$sum_metrics
```

## Search using various identifiers, including pubmed id, pmc id, and mendeley id

A single PubMed ID (pmid)

```{r}
alm_ids(pmid=22590526)
```

A single PubMed Central ID (pmcid)

```{r}
alm_ids(pmcid=212692)
```

## Search on many identifiers

```{r}
dois <- c('10.1371/journal.pone.0001543','10.1371/journal.pone.0040117',
	'10.1371/journal.pone.0029797','10.1371/journal.pone.0039395')
out <- alm_ids(doi=dois)
lapply(out$data, head)
```

## Get altmetrics by year

You can also get metrics by day (`sum_metrics='day'`) or month (`sum_metrics='month'`)

```{r}
alm_ids(doi='10.1371/journal.pone.0036240', sum_metrics='year')
```

## Output an-easy-to-combine-with-other-results data.frame

```{r}
alm_ids(doi='10.1371/journal.pone.0035869', total_details=TRUE)$data[,3:7]
```

## Get detailed data for altmetrics using `almevents`

```{r}
out <- alm_events(doi="10.1371/journal.pone.0029797")
names(out)
```

Get the results for PubMed Central

```{r}
# remove those with no data
out <- out[!out %in% c("sorry, no events content yet","parser not written yet")]
out[["pmc"]]
```

Get the results for twitter

```{r}
out$twitter$events[1:3,]
```

## Retrieve and plot PLOS article-level metrics signposts.

```{r fig.keep='high'}
dat <- alm_signposts(doi="10.1371/journal.pone.0029797")
plot_signposts(input=dat)
```

Or plot many identifiers gives a line chart

```{r fig.width=7, tidy=FALSE}
library("ggplot2")
dois <- c('10.1371/journal.pone.0001543','10.1371/journal.pone.0040117','10.1371/journal.pone.0029797','10.1371/journal.pone.0039395')
dat <- alm_signposts(doi=dois)
plot_signposts(input=dat) + 
  theme_grey(base_size = 12)
```

Or make an interactive chart by doing `plot_signposts(input=dat, type="multiBarChart")`. Try it out! It should open in your browser and you can interact with it.

## Density and histogram plots from PLOS Article Level Metrics data

Note: Do you the key below in the `searchplos` call in this example, but if you plan to use rplos more, get your own API key [here](http://api.plos.org/).

```{r}
library('rplos')
library('plyr')
dois <- searchplos(q='science', fl="id", fq=list('cross_published_journal_key:PLoSONE', 'doc_type:full', 'publication_date:[2010-01-01T00:00:00Z TO 2010-12-31T23:59:59Z]'), limit=75)
```

Remove non-full article DOIs

```{r}
dois <- dois$data$id
dois <- dois[!grepl("annotation", dois)]
```

Collect altmetrics data and combine to a `data.frame` with `ldply`

```{r}
alm <- alm_ids(doi=dois, total_details=TRUE)
alm <- ldply(alm$data)
```

The default plot

```{r fig.width=5, fig.height=4}
plot_density(alm)
```

You can change the color of the density plot

```{r fig.width=5, fig.height=4}
plot_density(alm, color="#EFA5A5")
```

Pass in a title or description subtending the title

```{r fig.width=5, fig.height=4}
plot_density(alm, title="Scopus citations from 2010")
```

Plot a particular source

```{r fig.width=5, fig.height=4}
plot_density(alm, source="crossref_total")
```

Plot many sources in different panels in the same plot, and pass in colors just for fun

```{r fig.width=7, fig.height=6, tidy=FALSE}
plot_density(input=alm, 
             source=c("counter_html","crossref_total","pubmed_total","scopus_total"),
             color=c("#83DFB4","#EFA5A5","#CFD470","#B2C9E4")) + 
  theme_grey(base_size = 12)
```

## CrossRef article level metrics

Notice that we are passing the base url for the Crossref API, whereas the default is for the PLOS url [http://alm.plos.org/api/v5/articles](http://alm.plos.org/api/v5/articles).

```{r}
url <- "http://det.labs.crossref.org/api/v5/articles"
dois <- c("10.1371/journal.pone.0086859", "10.1038/nature12990", "10.5860/choice.51-3037")
alm_ids(doi=dois, api_url = url)
```
