<!--
%\VignetteEngine{knitr}
%\VignetteIndexEntry{An R Markdown Vignette made with knitr}
-->

alm tutorial
======

## What are article level metrics? 

Glad you asked. The canonical URL for this is perhaps [altmetrics.org](http://altmetrics.org/manifesto/). Basically it is a metric that measures something about an article. This is in stark contrast to journal level metrics, like the Journal Impact Factor. 

## Are there other altmetrics data providers?

Yes indeedy. 

+ [ImpactStory](http://impactstory.it/)
+ [Altmetric.com](http://altmetric.com/)
+ [PlumAnalytics](http://www.plumanalytics.com/)

## Authentication

You aren't currently not required to use an API key to access the PLoS ALM API, but soon will need to.

Get your PLoS API key [here](http://api.plos.org/)

Put your API key in your .Rprofile file using exactly this: 
options(PlosApiKey = "YOUalmAPIKEY"), 
and the functions within this package will be able to use your API key without you having to enter it every time you run a search. 

## Install and load

You can get this package by installing via `install_github()` within Hadley Wickham's devtools package.

```{r install, eval=FALSE}
install.packages("devtools")
require(devtools)
install_github("alm", "rOpenSci")
```

```{r load, message=FALSE, warning=FALSE}
library(alm)
```

## The default call with either doi, pmid, pmcid, or mdid without specifying an argument for info

(We'll not print a few columns so the table prints nicely)

```{r alm1, message=FALSE, warning=FALSE, comment=NA, cache=TRUE}
alm(doi="10.1371/journal.pone.0029797")[,-c(6:8)]
```


## Details for a single DOI

```{r alm2, message=FALSE, warning=FALSE, comment=NA, cache=TRUE}
out <- alm(doi='10.1371/journal.pone.0029797', info='detail')
## totals
out[["totals"]][,-c(6:8)]
## history
head(out[["history"]])
```

## Search using various identifiers, including pubmed id, pmc id, and mendeley id

```{r alm3, message=FALSE, warning=FALSE, comment=NA, cache=TRUE}
# A single PubMed ID (pmid)
alm(pmid=22590526)[,-c(6:8)]
 
# A single PubMed Central ID (pmcid)
alm(pmcid=212692)[,-c(6:8)]
 
# A single Mendeley UUID (mdid)
alm(mdid="35791700-6d00-11df-a2b2-0026b95e3eb7")[,-c(6:8)]
```


## Search on many identifiers

```{r xxx, message=FALSE, warning=FALSE, comment=NA, cache=TRUE}
dois <- c('10.1371/journal.pone.0001543','10.1371/journal.pone.0040117',
	'10.1371/journal.pone.0029797','10.1371/journal.pone.0039395')
out <- alm(doi=dois)
lapply(out, head)
```

## Get altmetrics by year

You can also get metrics by day (`sum_metrics='day'`) or month (`sum_metrics='month'`)

```{r byyear, message=FALSE, warning=FALSE, comment=NA, cache=TRUE}
alm(doi='10.1371/journal.pone.0036240', sum_metrics='year')[,-c(6:8)]
```

## Output an-easy-to-combine-with-other-results data.frame

```{r easytocombine, message=FALSE, warning=FALSE, comment=NA, cache=TRUE}
alm(doi='10.1371/journal.pone.0035869', total_details=TRUE)[,3:10]
```

## Get altmetrics data for a single paper, and visualize the total data across dates

```{r totalmets, message=FALSE, warning=FALSE, comment=NA, cache=TRUE}
out <- alm(doi='10.1371/journal.pone.0001543', info='detail')
almplot(out, type='totalmetrics')
```

## Get detailed data for altmetrics using `almevents`

```{r almevents, message=FALSE, warning=FALSE, comment=NA, cache=TRUE}
out <- almevents(doi="10.1371/journal.pone.0029797")
names(out) # names of sources
out <- out[!out %in% c("sorry, no events content yet","parser not written yet")] # remove those with no data
out[["pmc"]] # get the results for PubMed Central
out[["twitter"]] # get the results for twitter (boo, there aren't any)
out[c("twitter","crossref")] # get the results for two sources
```

## Retrieve and plot PLOS article-level metrics signposts.

```{r signposts1, message=FALSE, warning=FALSE, comment=NA, cache=TRUE}
dat <- signposts(doi="10.1371/journal.pone.0029797")
plot_signposts(input=dat)
```

Or plot many identifiers gives a line chart

```{r signposts2, message=FALSE, warning=FALSE, comment=NA, cache=TRUE}
dois <- c('10.1371/journal.pone.0001543','10.1371/journal.pone.0040117','10.1371/journal.pone.0029797','10.1371/journal.pone.0039395')
dat <- signposts(doi=dois)
plot_signposts(input=dat)
```

Or make an interactive chart by doing `plot_signposts(input=dat, type="multiBarChart")`. Try it out! It should open in your browser and you can interact with it.

## Density and histogram plots from PLOS Article Level Metrics data

Note: Do you the key below in the `searchplos` call in this example, but if you plan to use rplos more, get your own API key [here](http://api.plos.org/).

```{r plot_densityprep, message=FALSE, warning=FALSE, comment=NA, cache=TRUE}
library(rplos); library(plyr)
dois <- searchplos(terms='*:*', fields="id", toquery=list('cross_published_journal_key:PLoSONE', 'doc_type:full', 'publication_date:[2010-01-01T00:00:00Z TO 2010-12-31T23:59:59Z]'), limit=100, key="WQcDSXml2VSWx3P")
alm <- alm(doi=do.call(c,dois$id), total_details=TRUE)
alm <- ldply(alm)
```

The default plot

```{r plot_densityplot1, message=FALSE, warning=FALSE, comment=NA, cache=TRUE}
plot_density(alm)
```

You can change the color of the density plot

```{r plot_densityplot2, message=FALSE, warning=FALSE, comment=NA, cache=TRUE}
plot_density(alm, color="#EFA5A5")
```

Pass in a title or description subtending the title

```{r plot_densityplot3, message=FALSE, warning=FALSE, comment=NA, cache=TRUE}
plot_density(alm, title="Scopus citations from 2010")
```

Plot a particular source

```{r plot_densityplot4, message=FALSE, warning=FALSE, comment=NA, cache=TRUE}
names(alm)[1:35]
plot_density(input=alm, source="crossref_citations")
```

Plot many sources in different panels in the same plot, and pass in colors just for fun

```{r plot_densityplot5, message=FALSE, warning=FALSE, comment=NA, cache=TRUE}
plot_density(input=alm, source=c("counter_total","crossref_citations","twitter_total","wos_citations"), color=c("#83DFB4","#EFA5A5","#CFD470","#B2C9E4"))
```

## Alt-metrics total citations from all sources.

```{r almtotals, message=FALSE, warning=FALSE, comment=NA, cache=TRUE}
almtotals(doi = '10.1371/journal.pbio.0000012')
```

## Get title of article by inputting the doi for the article.

```{r almtitle, message=FALSE, warning=FALSE, comment=NA, cache=TRUE}
almtitle(doi='10.1371/journal.pbio.0000012')
```